\documentclass[12pt]{amsart}

\usepackage{xcolor}
\usepackage[hidelinks]{hyperref}
\usepackage[margin=1.1in, tmargin=1in, bmargin=1in]{geometry}

\pagestyle{myheadings}
%\markleft{Syllabus for Math 331, Spring 2020}
%\markright{Syllabus for Math 331, Spring 2020}

\newcommand{\spacer}{\vspace{.2cm}}
\newcommand{\svs}{\vspace{.1cm}}

\newcommand{\red}[1]{\textcolor{red}{#1}}
\definecolor{gold}{rgb}{0.80,0.68,0.00}\newcommand{\gold}[1]{\textcolor{gold}{#1}}

\begin{document}

\begin{center}
\Large{\textsc{Syllabus}}
\end{center}

\noindent\hrulefill

\noindent\textbf{\textsc{Math 307 Linear Algebra and Diff. Eq.}} \hfill \textsc{Summer 2022}\\
Lecture: MWF 10:30am--11:55am \hfill Keller 404

\noindent\hrulefill

\spacer

\noindent\textbf{Instructor:} Pierre-Olivier Parisé (email: \texttt{parisepo@hawaii.edu})\\
Office: Physical Science Building (PSB) 302\\
Office hours: Email me to make an appointment.

\section*{Course description}
This course is intended to introduce the student to the fundamentals of linear algebra with an eye toward application of eigenvalue techniques to the solution of differential equations. The content to be covered includes matrices and Gaussian elimination, vector spaces, orthogonality, determinants, eigenvalues and eigenvectors, and differential equations.

\section*{Important Dates}
%Please make sure you write down somewhere these important dates:
Refer to the schedule on the course website for a detailed schedule.
	\begin{itemize}
	\item Exams:
		\begin{itemize}
		\item Midterm 1 (in-class): Friday, July 1st, 10:30am--12:00am.
		\item Midterm 2 (in-class): Friday, July 22nd 10:30am--12:00am.
		\item Final (in-class): Friday, August 12th, 10:30am--12:00am.
		\end{itemize}
	\item Non-instructional day(s):
		\begin{itemize}
		\item King Kamehameha Day: Friday, June 10th.
		\end{itemize}
	\end{itemize}

\section*{Required course material}

\noindent{\bf Textbook:} \emph{Linear Algebra and Differential Equations}, by Gary L. Peterson and James S. Sochacki. %Access to the textbook is required as some homework questions will be assigned out of it. I recommend forming study groups and sharing a textbook. 

\noindent{\bf Course website:} \url{https://mathopo.ca/courses-website/MATH-307/MATH-307.html}. Schedule, solutions to homework, and lecture notes are posted on the course website.



\section*{Grading components}
Your final grade will be calculated based on a weigthed average of the following components.
%You will be required to scan and upload your solutions to the homework on Gradescope (\texttt{http://gradescope.com}, entry code N87KN8).
\begin{enumerate}
\item{\bf Homework (40\%):} There will be homework each week assigned on each Mondays (starting on June 6th) and due for the next Monday, before midnight. The problems will be posted on Laulima under the assignment tab and the solutions will be posted in the Homework webpage of the course website. You may hand out your solutions during the Monday's lectures, send them to me by email, or put it in the envelop on my office door.
\item {\bf Midterm exams (40\%):} There will be two, closed notes, closed book, in-class midterms. Each exam will have a duration of 90min.
\item {\bf Final exam (20\%):} There will be a closed notes, closed book, in-class final.
\end{enumerate}

%\section*{Lectures}
%There will be lectures each Mondays, Wednesdays, and Fridays, 10:30am--11:55am taking place at Keller Hall 404. All students are expected to attend and participate in each lecture and take their own personal notes.

%If you miss a lecture, you are responsible for any assignments and/or announcements made. Unavoidable absences should be explained to the instructor. Office hours will not be utilized to re-teach material presented in a class missed. It is your responsibility to stay up-to-date.  

\section*{Missed assignment policies}

\noindent{\bf Policies for exams:} Attendance on the exams is compulsory; otherwise, a grade of zero will be recorded. Any student who has an excused, documented conflict with a test time must inform their instructor as soon as possible. Late requests will either be denied or will result in an automatic deduction from the exam score. An absence due to a positive COVID-19 test result must be notified to the instructor as soon as possible. For those students with an excused absence for a midterm, there will be a make-up exam.

Conflicts arising from work or social obligations, or from personal travel plans do \textbf{not} qualify as excused absences. By registering for this course, you are agreeing to take all exams at the scheduled times.

\noindent{\bf Policies for homework:} Late homework won't be accepted and will result of a mark of zero(0). If the guideline for a homework is not followed, at most 5\% will be retrieved from the homework total points. The guideline is available in the Homework webpage of the course website.
\svs

\noindent{\bf Academic integrity:}
All students are expected to abide by the university's Conduct Code. Academic sanctions for dishonesty may include receiving an F in the assignment or receiving an F in the class. There may be additional administrative sanctions.

\section*{Classroom policies}
Tablets can be used for note-taking only. During lectures, refrain from using electronic items, cell phones, music players, tablets, and laptops for any other purposes than note-taking.

Please arrive, be seated and ready to start each class on time. If you have a valid reason to leave early, please advise me before the class and try to sit near the exit to minimize disruption.

\section*{Sources of help}
%All students are  encouraged to come to \textit{office hours} to discuss homework questions or material from class. If the advertised times do not suit you, please email me for an appointment. It will be our pleasure to accomodate you for either in-person or online meeting.

{\bf KOKUA:} I am happy to work with you and the KOKUA Program (Office for Students with Disabilities), if you need course accommodations due to a disability. KOKUA can be reached at (808) 956-7511 or (808) 956-7612 (voice/text) in room 013 of the Queen Lili`uokalani Center for Student Services. All course modifications must be arranged through KOKUA. You are encouraged to start this process as early as possible.

\section*{Concerns}
If at any time during the semester you have any questions or concerns about the class, please contact me during regularly scheduled office hours or via email to make an appointment. You may also contact the following people:
\spacer

\noindent {\bf Director of Undergraduate Studies}\\
Mirjana Jovovic \\
Email: \texttt{undergrad-dir@math.hawaii.edu}

\svs
\noindent {\bf Associate Chair}\\
Bj{\o}rn Kjos-Hanssen \\
Email: \texttt{assoc-chair@math.hawaii.edu}

\end{document}

